Associative Data Structures for Website Filtering

### 2014-04-17 paper/paper.tex
Rewrote data sources and measurements.

    In technical terms, we are interested in the best data structure for associating many URLs with
    their respective permissions, be it whitelisted or blacklisted for certain users.  We have
    chosen to compare the performance of the binary search tree, the hash map, and the skip list, as
    these are commonly employed in practice.  It is of particular interest how each data structure
    will scale as the number of elements rises.  We have thus chosen for our main research question
    the following:
    % Guys, did you notice we did absolutely nothing about predicting this?

@@

@@

@@

    Using C++ as the implementation language, we chose libstdc++'s implementations of the standard class
    templates \texttt{std::map} and \texttt{std::unordered\_map} for the binary search
    tree and the hash table respectively.  As the C++ standard library does not contain a skip
    list\footnote{reference to the standard here}, we used an open-source third party library,
    \texttt{CSSkipList}\footnote{reference to library home page here}.  We also used libstdc++'s
    implementation of \texttt{std::string} for our string handling\footnote{specify?}needs.

    \subsection{Data Sources}
    For our data, we used datasets containing $2^x$ elements with $x$ an integer such that $8 < x <
    21$. Thus we have twelve different dataset sizes. We took the system's dictionary file and the
    list of top level domains from Wikipedia and combined these to generate a list of items in the
    form \texttt{www.\$word.\$tld}, where \texttt{\$word} is one of words from the the dictionary
    and \texttt{\$tld} is a top level domain.  Only words consisting of the lowercase letters
    \texttt{a} through \texttt{z} were used. Constraining the words to consist of 8 to 12 characters
    and limiting the top-level domains to exactly two letters, all possible combinations were
    generated.  For the full paths, a number of domains were selected and each was appended a random
    alphanumeric string of eighty characters.  Finally, for the third type of key, an IP address, we
    used 32 bit integers. The exact implementation can be viewed in the sourcecode\footnote{ref to
    apendix or code dump}.

    \subsection{Measurements}
    \label{subsec:measurements}
    For our time measurements, we used libstdc++'s implementation of the standard
    \texttt{std::chrono::high\_resolution\_clock} class.  There is no official documentation about
    the resolution of this clock, but measurements show that it is likely below one microsecond,
    which is sufficient for our purposes.\footnote{\url{http://stackoverflow.com/a/5524138/559931}}

    Memory measurements were performed by providing a custom allocator which tracked the number of
    memory requests.  Note that the latter deviates from our initial plan, where we
    stated\footnote{ref?}we would use the POSIX API\footnote{ref to the posix api}.  Unfortunately,
    that turned out to be too coarse-grained for our purposes and we could thus get no accurate
    measurements.  The allocator-based technique we employed did require some slight modifications
    to the skip list implementation we used, but this should not have caused a significant slowdown.
    We assume that any influence will only have made the tests more fair, as now allocation was done
    the same way in all three setups.

    After some tests we chose to decrease the number of runs used for measurements.  In the proposal
    we stated that for each time measurement we would perform the relevant operation 1000 times and
    then take the average of that.  In practice, this made the tests run for longer than we could
    afford. Because the results given by performing 100 runs were comparable, we settled for that.
    We expect the general result to be similar; refer to section\footnote{ref}for details about what
    \emph{could} have gone wrong.

>>>>

    In technical terms, we are interested in the best data structure for associating many URLs with
    their respective permissions, be it whitelisted or blacklisted for certain users.  We have
    chosen to compare the performance of the binary search tree, the hash table, and the skip list, as
    these are commonly employed in practice.  It is of particular interest how each data structure
    will scale as the number of elements rises.  We have thus chosen for our main research question
    the following:
    % Guys, did you notice we did absolutely nothing about predicting this?

@@

    For memory measurements, such repeated runs were deemed unnecessary as the deviation there was
    likely to be lower; a single map was thus constructed, and the difference in memory usage as
    shown by the POSIX API would be recorded.  As discussed in Subsection~\ref{subsec:measurements},
    this turned out to be unfeasible.

@@

    The following is an overview of our implementation.  More specific details can be found in the
    source code.\footnote{cite code here}

@@

    With C++ as the implementation language, we chose libstdc++'s implementations of the standard
    class templates \texttt{std::map} and \texttt{std::unordered\_map} for the binary search tree
    and the hash table respectively.  As the C++ standard library does not contain a skip
    list\footnote{reference to the standard here}, we used an open-source third party library,
    \texttt{CSSkipList}\footnote{reference to library home page here}.  We also used libstdc++'s
    implementation of \texttt{std::string} for our string handling\footnote{specify?}needs.

    \subsection{Data Sources}

    We performed tests using subsets of twelve different sizes, all of the form $2^x$ and with $8 <
    x \le 20$.  The word list was obtained by taking the system's dictionary file and selecting all
    words that contained only lowercase letters and that were at least 8 and at most 12 letters
    long.  The top-level domain list was obtained from Wikipedia, and filtered down to only those
    domains that consisted of two letters from the English alphabet.  The resulting domains were of
    the form \texttt{www.\$word.\$tld}, where \texttt{\$word} is one of words from the the word list
    and \texttt{\$tld} is a top level domain.  For the full paths, a number of domains were selected
    and each was appended a random alphanumeric string of eighty characters.  By using the same
    domain with multiple random strings, a sufficiently large number of paths was generated.
    Finally, as described in the research plan, consecutive 32-bit integers starting from zero were used
    for the IP addresses.

    \subsection{Measurements}
    \label{subsec:measurements}

    For our time measurements, we used libstdc++'s implementation of the standard
    \texttt{std::chrono::high\_resolution\_clock} class.  There is no official documentation about
    the resolution of this clock, but measurements show that it is likely below one microsecond,
    which is sufficient for our purposes.\footnote{\url{http://stackoverflow.com/a/5524138/559931}}

    Memory measurements were performed by providing a custom allocator which tracked the number of
    memory requests.  Note that the latter deviates from our initial plan, where we stated we would
    use the POSIX API.  Unfortunately, the measurements provided by that turned out to be too
    coarse-grained for our purposes, and our initial measurements were unusable.  The
    allocator-based technique we employed did require some slight modifications to the skip list
    implementation we used, but this should not have caused a significant slowdown.  We assume that
    any influence will only have made the tests more fair, as now allocation was done the same way
    in all three structures.

    After running a number of tests we chose to decrease the number of times we repeated each run.
    We had set out to perform each measurement 1000 times and take the average over that.  In
    parctice, however, the results obtained by performing only 100 runs did not vary greatly, and
    allowed us to perform more individual measurements.
